---
title: "Adavance r lab 6"
authors: "Hussnain & Arfan - Group-16"
date: "2022-10-18"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We imports foreach and doParallel libraries for parallel programming in brute_force_knapsack function

```{r setup}
library(foreach)
library(doParallel)
library(parallel)
library(profvis)
```

## Section 1: Brute force knapsack 
The brute_force_knapsack function used for finding all the possible combinations of max value per weight ratio of given data.

```{r}
brute_force_knapsack <- function(x, W, parallel = FALSE) {
  stopifnot(is.data.frame(x))
  stopifnot(W > 0)

  data <- x
  lengthOfCombinations <- 2^nrow(df)
  df <- data.frame()
  if (parallel == FALSE) {
    df <- foreach (i=1:nrow(data), .combine = rbind) %do% {
      elementCombinations <- combn(rownames(data), i, simplify = FALSE)
      combinedWeight <- colSums(combn(data$w, i))
      combinedValue <- colSums(combn(data$v, i))
      newDF <- data.frame(combinedWeight, combinedValue, I(elementCombinations))
      newDF
    }
  }else {
    numCores = detectCores()
    registerDoParallel(makeCluster(numCores))
    df <- foreach (i=1:nrow(data), .combine = rbind) %dopar% {
      elementCombinations <- combn(rownames(data), i, simplify = FALSE)
      combinedWeight <- colSums(combn(data$w, i))
      combinedValue <- colSums(combn(data$v, i))
      newDF <- data.frame(combinedWeight, combinedValue, I(elementCombinations))
      newDF
    }
    stopImplicitCluster()
  }
  df <- df[order(df$combinedValue, decreasing = TRUE),]
  left_weight <- W
  optimal_value <- 0
  elemets <- c()
  starts <- rep(100, 40)
  for (i in 1:nrow(df)) {
    row <- df[i,]
    if (left_weight > row$combinedWeight) {
      optimal_value <- optimal_value + row$combinedValue
      left_weight <- left_weight - row$combinedWeight
      elemets <- c(elemets, unlist(row$elementCombinations))
    }
  }
  return(list("value"=round(optimal_value), "elements" = as.numeric(elemets)))
}
```

We used below sample code to check our function results.

```{r}
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
data.frame(
w=sample(1:4000, size = n, replace = TRUE),
v=runif(n = n, 0, 10000)
)
```

We called our function using below values without parallel on and get expected results.

```{r}
brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500, parallel = FALSE)
```

Question How much time does it takes to run the algorithm for n = 16 objects?

The answers without parallel being true is: 
```{r}
system.time({
  brute_force_knapsack(x = knapsack_objects[1:16,], W = 3500, parallel = FALSE)
})
```

We than called our function with parallel being true and get expected results.

```{r}
# $value
# [1] 16770
# 
# $elements
# [1] 5 8
```

Question How much time does it takes to run the algorithm for n = 16 objects?

The answers with parallel being true is: 
```{r}
  #  user  system elapsed 
  # 1.373   0.024   2.284 
```

## Section 2: Knapsack dynamic
knapsack_dynamic problems is used for finding all the possible combinations of given data dynamically.

```{r}
knapsack_dynamic <- function(x, W) {
  data <- x
  mat <- matrix(nrow=nrow(data), ncol=W)
  k_recursive <- function(n, C) {
    if (isTRUE(mat[n,C])) {
      return(mat[n,C])
    }
    res <- 0
    if (n == 0 | C == 0) {
      res <- 0
    }else if (data$w[n] > C) {
      res <- k_recursive(n-1, C)
    }else {
      temp1 <- k_recursive(n-1, C)
      temp2 <- data$v[n] + k_recursive(n-1, C - data$w[n])
      res <- max(temp1, temp2)
    }
    mat[n,C] <- res
    return(res)
  }
  return(list(value = k_recursive(n = nrow(data), C = W)))
}
```

We called our function using below values and get expected results.

```{r}
knapsack_dynamic(x = knapsack_objects[1:8,], W = 3500)
```

Question How much time does it takes to run the algorithm for n = 500 objects?

The answer is:
```{r}
#   user  system elapsed 
#   160.756   43.323  203.069 
```

## Section 3: Greddy knapsack
greedy_knapsack problems is used for finding the best value over weight ratio of given data to the capacity of knapsack.

```{r}
greedy_knapsack <- function(x, W) {
  stopifnot(is.data.frame(x))
  stopifnot(W > 0)
  df <- x
  max_values <- df$v / df$w
  df$maxValue <- max_values
  rn <- rownames(df)
  df$elements <- rn
  df <- df[order(df$maxValue, decreasing = TRUE),]
  rownames(df) <- rn
  left_weight <- W
  optimal_value <- 0
  elements <- c()
  index <- 1
  repeat {
    if (left_weight >= df$w[index]) {
      optimal_value <- optimal_value + df$v[index]
      left_weight <- left_weight - df$w[index]
      elements <- c(elements, df$elements[index])
    }else {
      break
    }
    index <- index + 1
  }
  return(list("value"=round(optimal_value), "elements" = as.numeric(elements)))
}
```

We called our function using below values and get expected results.

```{r}
greedy_knapsack(x = knapsack_objects[1:800,], W = 3500)
```

Question How much time does it takes to run the algorithm for n = 1000000 objects?

The answers is:
```{r}
system.time({
  greedy_knapsack(x = knapsack_objects[1:1000000,], W = 3500)
})
```

## Section 4.1: Profiling
Profiling our functions code using profvis package:

```{r}
profvis::profvis({
  brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500, parallel = FALSE)
  knapsack_dynamic(x = knapsack_objects[1:8,], W = 3500)
  greedy_knapsack(x = knapsack_objects[1:800,], W = 3500)
})
```

## Section 4.2: Optimizing

After profiling with different results we find the solution of "memoization" to optimize our code. We implemented this technique in our knapsack_dynamic function. After implemeation of "memoization" our function memorize the previous results and instead of calculating again it just gets the stored data from a matrix that we created. Hence reduced the overall function time.

```{r}
knapsack_dynamic <- function(x, W) {
  data <- x
  mat <- matrix(nrow=nrow(data), ncol=W) # Memoization: Create a matrix for number of row and column as data
  k_recursive <- function(n, C) {
    if (isTRUE(mat[n,C])) { # Memoization: Checking if is already calculate and stored
      return(mat[n,C])
    }
    res <- 0
    if (n == 0 | C == 0) {
      res <- 0
    }else if (data$w[n] > C) {
      res <- k_recursive(n-1, C)
    }else {
      temp1 <- k_recursive(n-1, C)
      temp2 <- data$v[n] + k_recursive(n-1, C - data$w[n])
      res <- max(temp1, temp2)
    }
    mat[n,C] <- res # Memoization: Storing calculated result in matrix
    return(res)
  }
  return(list(value = k_recursive(n = nrow(data), C = W)))
}
```

